A Hedonic Price Analysis of Traffic Noise in the Twin Cities Housing Market

### 2013-06-26 reports/Draft.tex
updated the draft in preparation for meeting with Tsegaye

\usepackage{alltt}
\usepackage[round]{natbib}
\usepackage{amsmath, graphicx}

\title{A Hedonic Price Analysis of Traffic Noise in the Twin Cities Housing Market}
\date{}
\begin{document}
\maketitle

\section{Introduction}

\begin{itemize}
\item Introduction to environmental amenities of a property
\item Outline how the impact of environmental quality on the value of a property is specific depending on locational factors.
\item Intro to traffic noise as an environmental good.
\item Brief description regarding the negative impacts of traffic noise (brief summaries of past studies on the topic)
\item Outline how traffic noise is measured – decibel levels on a logarithmic scale
\item Description of the housing market crash in the fall of 2008
\item How the crash effected Twin Cities metro real estate market
\item Research question: Does traffic noise have a significant impact on properties within the Twin Cities’ metro real-estate market and if so, does this effect vary across space while accounting for time?
\end{itemize}

\section{Past Research}\label{sec:lit}
The value of noise pollution is typically estimated through simulated or related markets because no explicit private market for noise abatement exists. \citet{Nelson1982} is perhaps the best known earliest review of 14 studies of the economic impact of traffic noise on property values in Canada and the United States. More recently, a number of reviews have been produced for European government agencies. \citet{Bateman2001} reviewed the literature for the Scottish Government while \citet{Navrud2002} provided a ``state of the art'' review for the European Commission Environment Directorate General.

The Hedonic Price Method is commonly used to value noise by examining real estate price fluctuations as a function of noise levels. Based on the canonical work by \citet{Rosen1974}, researchers commonly estimate a hedonic price function specifying the logged sale price of a property as a linear function of the traffic noise exhibited at the house. Such a specification leads to reporting the impact of traffic noise as a percentage decrease in house price per one decibal (dB) increase in traffic noise. According to \citet{Batemen2001}, the negative effect of traffic noise has ranged from 0.08\% to 2.22\% per dB with a mean around 0.55 percent per dB. \citet{Navrud2002} found an average decrease of 0.64\% per dB in a review of 65 noise evaluation studies.

\citet{Nelson2008} produced an updated review of the traffic noise literature for the edited volume \underline{Hedonic Methods in Housing Markets Economics}. Nelson describes several areas of continued interest in the empirical literature, most notably: spatial modeling of the hedonic price function, long-run vs. short-run housing dynamics, and the appropriate measure of traffic noise. We now explain each of these areas in turn and our work's contribution to the field.

@@

>>>>

\usepackage{alltt}
\usepackage[round]{natbib}
\usepackage{caption, amsmath, graphicx}

\title{A Hedonic Price Analysis of Traffic Noise in the Twin Cities Housing Market}
\date{}

\begin{document}
\maketitle

\begin{abstract}
This paper implements the hedonic price method to estimate the impact of traffic noise on properties within the St. Paul, Minnesota metropolitan real-estate market over the time period from 2005 - 2010. We use Locally Weighted Regression (LWR) estimation techniques in order to allow the hedonic function to vary over time and space. Significant evidence suggests spatial and temporal non-stationarity in the hedonic price function within our study area. In contrast to previous research, we find little evidence that the impact of traffic noise exhibits a significant non-linear relationship with the decibel level of the traffic noise, nor does the impact seem to vary over space. Our results suggest that an increase in traffic noise of one decibel decreases house prices by an average of 0.19 percent before September, 2008, and an average of 0.37 percent after September, 2008. Both of these estimates are within the range estimated by previous research on the topic in other areas.
\end{abstract}

\section{Introduction and Past Research}\label{sec:lit}
The value of noise pollution is typically estimated through simulated or related markets because no explicit private market for noise abatement exists. \citet{Nelson1982} is perhaps the best known earliest review of 14 studies of the economic impact of traffic noise on property values in Canada and the United States. More recently, a number of reviews have been produced for European government agencies. \citet{Bateman2001} reviewed the literature for the Scottish Government while \citet{Navrud2002} provided a ``state of the art'' review for the European Commission Environment Directorate General.

The Hedonic Price Method is commonly used to value noise by examining real estate price fluctuations as a function of noise levels. Based on the canonical work by \citet{Rosen1974}, researchers typically estimate a hedonic price function specifying the logged sale price of a property as a linear function of the traffic noise exhibited at the house. Such a specification leads to reporting the impact of traffic noise as a percentage decrease in house price per one decibel (dB) increase in traffic noise. According to \citet{Bateman2001}, the negative effect of traffic noise has ranged from 0.08\% to 2.22\% per dB with a mean around 0.55\% per dB. In a review of 65 noise evaluation studies \citet{Navrud2002} found an average decrease of 0.64\% per dB.

\citet{Nelson2008} produced an updated review of the traffic noise literature for the edited volume \underline{Hedonic Methods in Housing Markets Economics}. The chapter describes several areas of continued interest in the empirical literature, most notably: the appropriate measure of traffic noise, spatial modeling of the hedonic price function, and housing market dynamics. We now explain each of these areas in turn and our work's contribution to the field.

Beyond the fact that no direct natural economic markets exist in which consumers can reveal their willingness to pay to avoid traffic noise, researchers also must choose how to measure the nuisance itself. Sound levels are commonly measured in decibels (dB), but traffic noise at a house can vary significantly depending on changing traffic conditions.

Some researchers (see, for instance, \citet{HughesJr.1992}) measure the negative externality of traffic by correlating traffic volume with house prices. However, traffic may generate other externalities beyond noise (such as local air pollution and reduced safety for pedestrians and drivers). Given the interest in government policymakers in traffic noise mitigation (cite some data here), we focus on the negative externality stemming specifically from the traffic noise itself. Additionally, until 2011, the Minnesota Department of Transportation used a cost effectiveness threshold for noise mitigation of \$3,250 per dB reduced. Given that some of the benefits of noise reductions are capitalized in house prices, we are curious to see how our estimated magnitudes compare to the state cost effectiveness cutoff.

I'm going to need some help writing about noise measures. Perhaps Tsegaye can lend a hand?

\citet{Huang;Palmquist2001} collect noise data using 128 monitoring stations throughout a suburb of Seattle, WA. They use dBA $L_{10}$ as their noise measure, where $L_{10}$ denotes the value that is exceeded 10 percent of the time. They collapse the noise readings into 2.5 dBA wide bins and report noise impact estimates of roughly 1.4\% decrease in house sales price per category increase in noise (or about 0.4\% per dBA $L_{10}$).

@@

% \begin{table}
% \caption{Data Description}\label{tab:descriptive}
% \begin{tabular}{lp{3in}r}
% Variable & Description  & Expected Hedonic Impact \\ \
% 
% Sale Price &   Price of most recent sale  (in \$) & \\
% 
% Home Size	& Area of finished space in the home (in square feet)	& (+) \\
% Lot Size	& Area of land house sits on (in acres)	& (+) \\
% Year Built & Year home was constructed &	(+) \\
% Owner	Occupancy & Does the owner occupy the house? (0 or 1) &	(+) \\
% 
% Traffic Noise	& Maximum decibel level of traffic noise (dB) measured within parcel &	(-) \\
% MCA 3  & Average 3rd grade Minnesota Comprehensive Assessment (MCA) score for local elementary school &	(+) \\
% Median Income	& Median household income (in dollars) for 2010 Census tract &	(+) \\
% Dist. to CBD &	Distance (in meters) to nearest central business district	& (-) \\
% Dist. to shop &	Distance (in meters) to nearest major shopping center	& (+)/(-) \\
% Dist. to lake &	Distance (in meters) to nearest public lake	& (-) \\
% Dist. to park	& Distance (in meters) to nearest public park	& (-) \\
% City &	City a parcel resides in	& (+)/(-) \\ \hline
% \multicolumn{3}{l}{Total number of house sales between 2005 and 2010 (n = 42,095) }
% \end{tabular}
% \end{table}
